\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

First, I would like to thank my supervisor, Prof. Steve Blackburn.
When I first came to Steve's lab, he did what he can to help me get familiar with
this project and the area of virtual machines and garbage collection.
Steve also gave me a lot of instructions with his wisdom and knowledge
and helped me solved all the problems I was facing throughout my research project this year.
Steve's knowledge and experience in the area of memory management and
programming language impressed me a lot.
Having academic discussions and debates with Steve was a great fun and
I would love to continue working with him in the future if possible.
In addition, Steve provided me with a lot of opportunities in the area of research,
like funding me for attending conferences, which strongly helped me to become a better researcher.

I would like to thank other members in our research lab, especially
Branda Wang, Zixian Cai and Pavel Zakopaylo, who gave me much of help
with the project,  especially when I first came to this lab and barely know nothing
about garbage collection and Java Virtual Machines.
Also, I want to thank Lin Yi and Xi Yang, although we have never met, your useful
instructions and advice about the benchmarks made my life much easier when I was
performing measurements for my project.

Also, I want to thank my partner Tianjia Zhao who gave me plenty of support throughout this year.
Conducting a one-year research project alongside with four hard computer science courses is no easy,
and you always try your best to comfort me every time when I was under stress.

Finally, I would like to thank my parents Houjian Zhao and Ying Qiang, 
for raising me and provided me with a good environment for my development.
Especially, you inspired my interest in science when I was a young child,
which I will never forget in the rest of my life.
Although we are in separate countries this year,
you always care about my life and provide me with plenty of support for my life and research.

