\chapter{Future Work}
\label{cha:future}

This chapter will discusses the future work I plan to do after the completion
of thie thesis. It mainly includes:
$1.$ Resolve a data rece problem for Shenandoah GC.
$2.$ Perform some optimizations
$3.$ Start wkring on ZGC.

\section{Race Problem for Shenandoah GC}

Currently the implementation of Shenandoah GC has a data race problem.
Based on the previous analysis, the mutator is not lock and release a monitor
expectedly during the concurrent phase of the Shenandoah GC, when there can be
two different copies of an object exist in the heap.

Duo to the time scope of thie thesis, I cannot solve this issue currently.
But I have implemented a work-around in the Shenandoah GC, with lower performance.
As the most urgent problem I are facing, I plan to solve this issue immediately after
this project.

\section{Optimizations}

I am also considering the possibility to make the implemented Garbage-first
and Shenandoah GC become production ready. One major part missing is the Optimizations
since the development of these garbage collectors was following the original design
of these collectors and did not have too much optimizations. After performing some
optimizations on these collectors as well as some additional correctness verfication,
the collector can have the possibility to become production ready.

\section{ZGC}

Due to the time scope of this project, I did not implement and measure ZGC.
In addition, since JikesRVM and MMTk only supports the 32bit address space but the pointer
coloring process in ZGC requires the 64bit address space, this makes the implementation
of ZGC more difficult.
But as a latest member of the Garbage-first family of collectors, ZGC is
worth for implementation and measurements.
However, a detailed plan of resolving several hardware incompactabilities should be done in the future,
before starting the implementation of ZGC.

\section{Summary}

This chapter will discusses the future work I plan to do after the completion
of thie thesis. The next chapter contains the conclusion I obtained from the
development and the analysis results of the Garbage-first family of garbage collectors.